\documentclass[a4paper,UKenglish,cleveref, autoref]{svjour3}
%\usepackage{latexsym,amsthm,amsmath,amssymb}


%\usepackage[utf8x]{inputenc}
%\usepackage{bm}
%\usepackage{pbox}
%\usepackage{type1ec}
\usepackage[T1]{fontenc}
\usepackage{bbm}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{pifont}
\usepackage{url}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{fancybox}
\usepackage{colortbl}
\usepackage{soul}
\usepackage{color}
\usepackage{tikz}
\usepackage{tkz-berge}
\usepackage{rotating}
\usepackage{pdflscape}
\usepackage{afterpage}
\usepackage{graphicx}
\usepackage{comment}
\usepackage{etoolbox}
\usepackage{environ}
\usepackage{dirtytalk}

%% Ignasi's packages %%
\usepackage{cite}
\usepackage{hyperref}

\usepackage{algorithm}
\usepackage{algorithmic}

\newcommand{\ig}[1]{\textcolor{red}{[Ig: #1]}}
\newcommand{\gcm}[1]{\textcolor{blue}{[Gui: #1]}}

\newcommand{\probl}[3]{
  \begin{flushleft}
    \fbox{
      \begin{minipage}{0.95\textwidth}
        \noindent {\sc #1}\\
        {\bf Instance:} #2\\
        {\bf Question:} #3
      \end{minipage}}
    \medskip
  \end{flushleft}
}

\newcommand{\paraprobl}[5]
{
  \begin{flushleft}
    \fbox{
      \begin{minipage}{#5cm}
        \noindent {\textsc {#1}}\\
        {\bf Instance:} #2\\
        {\bf Parameter:} #4\\
        {\bf Question:} #3
      \end{minipage}
    }
  \end{flushleft}
}

%\newcommand{\paraprobl}[4]
%{
%  \begin{flushleft}
%    \fbox{
%      \begin{minipage}{14cm}
%        \noindent {\textsc {#1}}\\
%        {\bf Input:} #2\\
%        {\bf Parameter:} #4\\
%        {\bf Output:} #3
%      \end{minipage}
%    }
%  \end{flushleft}
%}

\newcommand{\paraprobldecision}[4]
{
  \begin{flushleft}
    \fbox{
      \begin{minipage}{.95\textwidth}
        \noindent {\textsc {#1}}\\
        {\bf Input:} #2\\
        {\bf Parameter:} #4\\
        {\bf Question:} #3
      \end{minipage}
    }
  \end{flushleft}
}


\newcommand{\Acal}{\mathcal{A}}



\usetikzlibrary{decorations,arrows,shapes}


\usepackage{complexity}


\newtheorem{rrule}{Reduction Rule}
\newtheorem{observation}{Observation}

\newcommand{\balpha}{\boldsymbol{\alpha}}
\newcommand{\bmd}{\boldsymbol{d}}

\newcommand{\bigO}[1]{\mathcal{O}\!\left(#1\right)}
\newcommand{\bigOs}[1]{\mathcal{O}^*\!\left(#1\right)}

\newcommand{\dcup}{\ \dot\cup\ }

\NewEnviron{sproof}[1] {
    \begin{proof}[Proof of safeness of Rule~#1]\BODY\end{proof}
}

\newcommand{\ceil}[1]{\left\lceil#1\right\rceil}
\newcommand{\floor}[1]{\left\lfloor#1\right\rfloor}

\newcommand{\pname}[1]{\textsc{#1}}

\newcommand{\nproblem}[3]{{\centering\fbox{\pbox{\textwidth}{\pname{#1}\\\textit{Instance}: #2\\\textit{Question}: #3}}}}
\newcommand{\pproblem}[4]{{\centering\fbox{\pbox{\textwidth}{\pname{#1}\\\textit{Instance}: #2\\\textit{Parameter}: #3\\\textit{Question}: #4}}}}

%tikz stuff
\newcommand{\inners}{1.2pt}
\newcommand{\outers}{1pt}
\newcommand{\gscale}{0.6}

\newcommand{\tdef}[1]{\emph{#1}}


%complexity
\newclass{\Hard}{hard}
%\newcommand{\Dificil}{\textsf{\normalfont Difícil}}
%\newcommand{\Dificuldade}{\textsf{\normalfont Dificuldade}}
\newclass{\Hness}{hardness}
\newcommand{\NPH}{\NP\text{-}\Hard}
\newclass{\Complete}{complete}
\newclass{\para}{para}
\newcommand{\pNPH}{\para\text{-}\NP\text{-}\Hard}
%\newclass{\Completo}{Completo}
%\newclass{\Ctude}{Completude}
\newclass{\Cness}{completeness}
\newcommand{\NPc}{\NP\text{-}\Complete}
\newcommand{\NPct}{\NP\text{-}\Ctude}
\newcommand{\NPcness}{\NP\text{-}\Cness}
\newcommand{\NPhness}{\NP\text{-}\Hness}
\newfunc{\tw}{tw}
\newfunc{\YES}{YES}
\newfunc{\NOi}{NO}
\newfunc{\dc}{dc}
\newfunc{\dcc}{d\overline{c}}
\newclass{\ETH}{ETH}
\newclass{\dgr}{deg}


\title{Finding Cuts of Bounded Degree\thanks{G.C.M. Gomes was funded by Coordenação de Aperfeiçoamento de Pessoal de Nível Superior - Brasil (CAPES) - Finance Code 001.
\\ I. Sau was funded by Projects DEMOGRAPH (ANR-16-CE40-0028) and ESIGMA (ANR-17-CE23-0010).}}
\subtitle{Complexity, FPT and Exact Algorithms, and Kernelization}

\author{Guilherme C. M. Gomes \and Ignasi Sau}

\institute{G. C. M. Gomes \at Universidade Federal de Minas Gerais, Departamento de Ciência da Computação, Belo Horizonte, Brazil.
          \and
          G. C. M. Gomes \at LIRMM, Universit\'e de Montpellier, Montpellier, France.
              \\\email{gcm.gomes@dcc.ufmg.br}           %  \\
%             \emph{Present address:} of F. Author  %  if needed
           \and
           I. Sau \at CNRS, LIRMM, Universit\'e de Montpellier, Montpellier, France.
              \\\email{ignasi.sau@lirmm.fr}
}

\begin{document}

\maketitle

\begin{abstract}
A \emph{matching cut} is a partition of the vertex set of a graph into two sets $A$ and $B$ such that each vertex has at most one neighbor in the other side of the cut. The \textsc{Matching Cut} problem asks whether a graph has a matching cut, and has been intensively studied in the literature. Motivated by a question posed by Komusiewicz et al.~[IPEC 2018], we introduce a natural generalization of this problem, which we call \textsc{$d$-Cut}: for a positive integer $d$, a \emph{$d$-cut} is a bipartition of the vertex set of a graph into two sets $A$ and $B$ such that each vertex has at most $d$ neighbors across the cut.  We generalize (and in some cases, improve) a number of results for the \textsc{Matching Cut} problem. Namely, we begin with an \NP-hardness reduction for \textsc{$d$-Cut} on $(2d+2)$-regular graphs and a polynomial algorithm for graphs of maximum degree at most $d+2$. The degree bound in the hardness result is unlikely to be improved, as it would disprove a long-standing conjecture in the context of internal partitions. We then give {\sf FPT} algorithms for several parameters: the maximum number of edges crossing the cut, treewidth, distance to cluster, and distance to co-cluster. In particular, the treewidth algorithm improves upon the running time of the best known algorithm for \textsc{Matching Cut}. Our main technical contribution, building on the techniques of Komusiewicz et al.~[IPEC 2018], is a polynomial kernel for \textsc{$d$-Cut} for every positive integer $d$, parameterized by the distance to a cluster graph. We also rule out the existence of polynomial kernels when parameterizing simultaneously by the number of edges crossing the cut, the treewidth, and the maximum degree. Finally, we provide an exact exponential algorithm slightly faster than the naive brute force approach running in time $\bigOs{2^n}$.
\keywords{matching cut; bounded degree cut; parameterized complexity; {\sf FPT} algorithm; polynomial kernel; distance to cluster.}
\end{abstract}



\input{sections/intro.tex}
\input{sections/np.tex}
\input{sections/param.tex}
\input{sections/conclusion.tex}


%\bibliographystyle{plainurl}
\bibliographystyle{abbrv}
\bibliography{refs}

\end{document}



